Capturing Policies for BYOD

### 2016-12-20 paper.tex
davids feedback

\usepackage{acronym}
\acrodef{MDM}{Mobile Device Management}
\acrodef{BES}{Blackberry Enterprise Services}
\acrodef{HiMSS}{Healthcare Information Management System Society}

\usepackage{listings}
\lstdefinelanguage{AppPAL}{%
  morekeywords={if,inf,says,where,true,false,with,is},
  otherkeywords={can-act-as,can-say},
  sensitive=true,
  morestring=[b]',
  literate={\ inf\ }{{$\infty$}}5
}[keywords,strings]
\newcommand{\listingsize}[0]{\footnotesize}
\lstset{%
  basicstyle=\normalfont\ttfamily\listingsize,
  keywordstyle=\normalfont\ttfamily\slshape\listingsize,
  stringstyle=\normalfont\sffamily\listingsize,
  columns=flexible,
  mathescape,
  tabsize=2,
  showstringspaces=true,
  %numbers=left,
  frame=single,
  breaklines=true,
  numberstyle=\tiny\ttfamily\color{Gray},
  commentstyle=\ttfamily\color{Gray}
}
\newcommand{\code}[2][]{\lstinline[breaklines=true, #1]{#2}}

\usepackage{mdframed}
\newcommand{\todo}[1]{\marginpar{\begin{mdframed}[userdefinedwidth=1cm]\tiny\sffamily\raggedright #1\end{mdframed}}}

\newenvironment{policyrule}[1]{%
  \begin{mdframed}\footnotesize
      \noindent\textbf{\sffamily #1}:~\itshape%
}{%
  \end{mdframed}
}

\title{Capturing Policies for BYOD}
\author{Joseph Hallett and David Aspinall}
\institute{University of Edinburgh}
\begin{document}
\maketitle
\begin{abstract}
  % 4 Sentences
  BYOD policies are informally specified using natural language.
  Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
  The \AppPAL~language is shown capturing BYOD policies throughout the paper.
  We identify common BYOD idioms and delegation structures.
\end{abstract}
\section{Introduction}
\label{sec:intro}
% Describe the problem
% State my contribution
% 1 page

Employees bring their own devices to work.
In the past employees might have had a dedicated company device.
Today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016} and, in some fields,
  85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
Controlling employee's devices is a challenge for IT departments.
They need to control access to company resources but have limited control over the devices used to access them.

One solution to controlling devices is to require users to agree to follow policies.
They often take the form of a user agreement, written in natural language, which describes how the devices should be used and configured.
Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
On top of user agreements, companies may also use \ac{MDM} software; which can enforce policies.
But the use of \ac{MDM} software does not guarantee compliance.
One survey found over 50\% of companies with \ac{MDM} software have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.

The policies are becoming more intricate.
\ac{MDM} software is becoming more proficient.
There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
Implementing a policy is only part of the problem, however.
BYOD policies are specified informally using natural language, and they contain more than just access control decisions.
These policies describe trust relationships inside the company between the IT departments, users, and HR each who may be delegated to make decisions and provide rules.
Policies contain rules that require employees to acknowledge risks, and regulations.
An antivirus or \ac{MDM} program may be used to \emph{implement} part of the policy.
But it is the policy, however, that \emph{specifies} which software to use and when. 
There is no automatic way to check how the policy has been implemented and by what.

Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
Our formalizations pick out the common concerns and trust relationships in these policies.
We look at what decisions and trust relationships used in BYOD policies.
These are BYOD idioms that are capture decisions and frequently seen requirements in BYOD policies.
These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.

We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
Our AppPAL implementation can also be easily extended to support new types of policies.

\section{Related Work}
\label{sec:related}
% 1-2 pages

@@

Martinelli~\etal{}'s looks at creating a dynamic permissions manager, called UC-Droid.
Their tool can alter what an app's Android permissions are at run time based on policies~\cite{martinelli_enhancing_2016}.
The tool allows companies to reconfigure the apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
These kinds of policies are more configurable than the geo-fenced based policies existing MDM packages provide.
Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
The work allowed a company to verify the devices within their network and what servers and services they could access.
It also describes a mechanism for providing different users with different policies.

Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
Their tool allows companies to distribute apps through a secure app~store.
The store ensures apps meet policies through a combination of static analysis and app rewriting with dynamic enforcement.
Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
Using their tool, they implemented parts of a NATO Communications and Information Agency policy relating to personal networks and data management~\cite{armando_developing_2016}.
Their work shows how the app-specific sections of a BYOD policy can be check and enforced using tools.
They did not look at where the checks or policies come from, however.

@@

Other work has looked at using app wrapping (where an app is recompiled to use guarded APIs) to enforce more general policies.
Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
App rewriting has the advantage that the device's underlying OS needn't be modified as the apps are changed at the source code level.
Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}. 
However app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.

@@

As mobile devices have become more common in the workplace, BYOD policies have grown increasingly complex
Part of their policies are prescriptive:  if you configure your device in this way, you will mitigate that threat.
The policies contain more than just configuration, however.
Consider this rule taken from the \emph{Security Policy for the use of handheld devices in corporate environments} by SANS~\cite{nicholas_r._c._guerin_security_2008}:

@@

A company could make use of an \ac{MDM} program, such as IBM's MaaS360 or \ac{BES}~\cite{_ibm_????,_secure_????} to implement the policy.
These programs support app wrapping, and security ``checkboxes'' to enable security controls and access control policies.
Geofencing could also be used to apply additional policies in the area around a lab.
Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration.
It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
Does it understand how the risk was analysed?
Can it judge using a policy if a location is restricted?
The rule also gives a security objective: \emph{prevent malicious users from sending proprietary pictures}.
The guidelines are given, however, for the case of a legitimate user using MMS or email.
It may not be sufficient to stop a sufficiently motivated \emph{malicious} user.

To answer these questions we present \AppPAL{}: 
  a tool for linking policies to the tools used to implement them and distributing decisions about apps.
Our approach does not try to enforce the policy by checking the app's code for programming errors.
Rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration are made explicit.
This gives us greater clarity as to which tool is being trusted to implement what policy.
It allows us to see who is being trusted to make which decisions,
  and use automatic-tools to uncover problematic aspects of the policy~\cite{hallett_specifying_2016}.
Continuing with the example above, we can write this in \AppPAL~as:

@@

\section{Changes from SecPAL}
\label{ssec:changes}

@@

Facts of the \emph{must}-kind represent obligations, actions to complete if a particular scenario presents itself.
For these facts, we add a rule to check the we perform the obligation.
This rule should be checked periodically to ensure compliance.
Our implementation contains tooling to generate these rules automatically.
\begin{lstlisting}
$\langle\text{speaker}\rangle$ says $\langle\text{subject}\rangle$ hasSatisfiedObligation$\langle\text{Action}\rangle$
  if $\langle\text{subject}\rangle$ must$\langle\text{Action}\rangle$,
     $\langle\text{subject}\rangle$ has$\langle\text{Action}\rangle$.
\end{lstlisting}

We often need to describe the type of variables.
Facts using \emph{is} predicates give types to variables; 
  however, \AppPAL~inherits from SecPAL's (and Datalog's) safety condition that the body of an assertion must reference all the variables in the head.
Following the rule leads to policies which can be noisy to read.
In one policy we looked at (SANS), there is a rule that states that devices can only connect to wireless access points owned by the company.
\begin{lstlisting}
'company' says Device canConnectToAP(AP)
  if AP isOwnedByCompany,
     Device isDevice,
     AP isAP.
\end{lstlisting}
To simplify the policies, we add syntactic sugar for statements giving variables their type (\emph{variable \emph{is}Type}).
Variables in the head of the rule of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the body.
To hide typing statements we rewrite the rule as shown below emphasising the policy rule rather than SecPAL boiler-plate.
\begin{lstlisting}
'company' says Device:D canConnectToAP(AP:X)
  if X isOwnedByCompany.
\end{lstlisting}

@@

We analysed five different policies looking for common idioms.
The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
This policy is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices.
Companies are expected to modify the document to suit their needs.
The policy is general; not specific to any particular industry, device, or country's legislation.
The second is taken from the \ac{HiMSS}~\cite{healthcare_information_and_management_systems_society_mobile_2012};
  a US non-profit company trying to improve healthcare through IT.
The \ac{HiMSS} policy is relatively short and contains concerns specific to healthcare scenarios. 
It is written as a contract the users agree to follow.
In contrast, every other policy we looked at is written as an organisation imposing rules on users they should follow to ensure compliance.
The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
The third is taken from a British hospital trust~\cite{kennington_mobiles_2014} and describes the BYOD scheme used in practice at the hospital.
Finally, we looked at two simpler policies from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
These policies are simpler than the other policies we looked at comprised of much more general rules and around two pages in length.

\begin{table}\centering\footnotesize\sffamily
  \resizebox{\textwidth}{!}{%
    \def\myfntx{}
    \begin{tabular}{l c c c c c}
      \toprule
                                      & {SANS}       & {HiMSS}     & {NHS}       & {Edinburgh}                       & {Sirens}    \\
      \midrule
      Number of rules                 & 33           & 15          & 56          & 20                                & 25          \\
      \AppPAL~assertions              & 71           & 21          & 58          & 10                                & 39          \\
      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large nmuber of rules that whilst marked as rules are infact just descriptions of the document.  All the policy rules that described restrictions or relationships were implemented in \AppPAL{}.} & 22 (88$\%$) \\
      \midrule
      Rules using Acknowledgement     & 2            & 10          & 11          & 1                                 & 6           \\
      Rules using Delegation          & 23           & 5           & 33          & 2                                 & 13          \\
      Rules describing a restriction  & 18           & 3           & 8           & 1                                 & 5           \\
      \midrule
      Principal Speaker               & company      & user        & nhs-trust   & records-management                & department  \\
    %Technical Speaker            & it-department & xyz-healh-system & it-department &                    & it-department \\
      \bottomrule                    \\
    \end{tabular}
  }
  \caption{Summary of the contents of each of the BYOD policies.}
  \label{tab:summary}
\end{table}

We summarise the policies in \autoref{tab:summary}.
Each policy contains a series of \emph{rules}, which we implemented by one or more \emph{\AppPAL~assertions}.
The \emph{policy coverage} represents the number of rules that have an \AppPAL~description attached.

Not all the rules in a policy require implementation.
For example rule 9.1 in the NHS policy states:
\begin{policyrule}{NHS}
  Downloading of personal apps onto a corporately issued mobile device should be avoided where possible.
  The Trust would not encourage staff members to download apps for personal use onto a corporately issued mobile device.
  All staff are reminded that they must adhere to the guidance outlined in the Social Media Policy.
\end{policyrule}
The rule does not prohibit any action and doesn't require the staff to do anything different.
It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
%\begin{lstlisting}
%'nhs-trust' says Staff:S mustAcknowledge('avoid-installing-personal-apps').
%'nhs-trust' says Staff:S mustAcknowledge('social-media-policy').
%\end{lstlisting}
In this case, however, since nothing is required even an acknowledgement seems excessive.

@@

\section{BYOD Idioms}

When examining the policies, we noticed two particular idioms reappearing: acknowledgements and delegation.
We describe both idioms in greater detail, and show how they can be implemented in \AppPAL{}, below.
MDM tools and research have focussed so far on implementing restrictions on apps and devices~\cite{_ibm_????,armando_formal_2014,martinelli_enhancing_2016}.
Implementing these controls is a vital aspect of BYOD policies and all 5 of the policies we looked at had rules that described restrictions~(\autoref{tab:summary}).
Every policy also contained rules that required employees acknowledgements, however.
Only the SANS policy (which is configuration focussed) contained more rules that required restrictions than acknowledgements. 
All the policies contained more rules featuring delegation relationships than functionality restrictions.
Restricting device functionality is tricky and important, but other aspects of BYOD policies are also worth attention.

@@

\AppPAL~requires an explicit speaker for each assertion.
Speakers can delegate to others by making a statement about what they \emph{can-say}.
When translating the policies, the author of the policy is used as the primary speaker of the policy's rule (\autoref{tab:summary}).
For the \ac{HiMSS} policy, where the user states what they will do rather than the company stating what they must, the user is the primary speaker.
All the policies describe multiple entities that might make assertions and delegate.
With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
The delegation might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.

@@

\section{Comparison with Existing MDM Software}

It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}, both leading MDM packages~\cite{rob_smith_magic_2016}.
Each of these tools supports enforcing and checking compliance policies. 
They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.

>>>>

\usepackage{acronym}
\acrodef{MDM}{Mobile Device Management}
\acrodef{BES}{Blackberry Enterprise Services}
\acrodef{HiMSS}{Healthcare Information Management System Society}
\acrodef{GSN}{Goal Structuring Notation}

\usepackage{listings}
\lstdefinelanguage{AppPAL}{%
  morekeywords={if,inf,says,where,true,false,with,is},
  otherkeywords={can-act-as,can-say},
  sensitive=true,
  morestring=[b]',
  literate={\ inf\ }{{$\infty$}}5
}[keywords,strings]
\newcommand{\listingsize}[0]{\footnotesize}
\lstset{%
  basicstyle=\normalfont\ttfamily\listingsize,
  keywordstyle=\normalfont\ttfamily\slshape\listingsize,
  stringstyle=\normalfont\sffamily\listingsize,
  language=AppPAL,
  columns=flexible,
  mathescape,
  tabsize=2,
  showstringspaces=true,
  %numbers=left,
  frame=single,
  breaklines=true,
  xleftmargin=1em,
  numberstyle=\tiny\ttfamily\color{Gray},
  commentstyle=\ttfamily\color{Gray}
}
\newcommand{\code}[2][]{\lstinline[breaklines=true, #1]{#2}}

\usepackage{mdframed}
\mdfdefinestyle{supertight}{%
  leftmargin=0pt,
  rightmargin=0pt,
  skipabove=1em,
  skipbelow=1em,
  innerleftmargin=2pt,
  innerrightmargin=2pt,
  innertopmargin=2pt,
  innerbottommargin=2pt,
}

\newcommand{\todo}[1]{\marginpar{\begin{mdframed}[style=supertight]\tiny\sffamily\raggedright #1\end{mdframed}}}

\newenvironment{policyrule}[1]{%
  \begin{mdframed}[]\footnotesize
      \noindent\textbf{\sffamily #1}:~\itshape%
}{%
  \end{mdframed}
}

\title{Capturing Policies for BYOD}
\author{Joseph Hallett and David Aspinall}
\institute{School of Informatics, University of Edinburgh}
\begin{document}
\maketitle
\begin{abstract}
  % 4 Sentences
  BYOD policies are informally specified using natural language.
  We show how the \AppPAL~language can help reduce ambiguity in BYOD policies and link the specification of a BYOD policy to its implementation.
  This paper identifies a BYOD idiom that is ignored by existing MDM tools.
  We explore the delegation relationships and concerns in 5 BYOD policies, and describe how \AppPAL might be used to implement them using existing tools.
\end{abstract}
\section{Introduction}
\label{sec:intro}
% Describe the problem
% State my contribution
% 1 page

Employees bring their own devices to work.
In the past employees might have had a dedicated company device.
Today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016} and, in some fields,
  85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
Controlling employee's devices is a challenge for IT departments.
Failure to do so can lead to cause employees to leak confidential information and the waste of company resources.
Unfortunately the companies have limited control over the devices used to access them.

One solution to controlling devices is to require users to agree to follow policies.
They often take the form of a user agreement, written in natural language, which describes how the devices should be used and configured.
Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
On top of user agreements, companies may also use \ac{MDM} software which can help enforce policies.
\ac{MDM} software can configure a device's security settings, and add 
But the use of \ac{MDM} software does not guarantee compliance.
One survey from a leading MDM vendor found over 50\% of companies with their \ac{MDM} software still had devices that do not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
Reasons for non-compliance included out-of-date policies and employees tampering with the \ac{MDM} software.

BYOD policies are becoming more intricate.
There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
Implementing a policy is only part of the problem, however.
BYOD policies are specified informally using natural language, and they contain more than just access control decisions.
These policies describe trust relationships inside the company between the IT departments, users, and HR each who may be delegated to make decisions and provide rules.
Policies contain rules that require employees to acknowledge risks, and regulations.
An antivirus or \ac{MDM} program may be used to \emph{implement} part of the policy.
But it is the policy, however, that \emph{specifies} which software to use and when. 
There is no automatic way to check how the policy has been implemented and by what.

To describe the policies we came up with \AppPAL{}: 
  a formal language for linking policies to the tools used to implement them and distributing decisions about apps.
Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
Our formalizations pick out the common concerns and trust relationships in these policies.
We look at what decisions and trust relationships used in BYOD policies.
These are BYOD idioms that are capture decisions and frequently seen requirements in BYOD policies.
These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.

@@

\AppPAL~is not designed to replace existing static and dynamic analysis and enforcement tools.
A company might use multiple different tools, app stores, and even contractual agreements with employees to enforce their policies.
We aim to help clarify the meaning of ambiguous natural language policy documents, and provide a rigorous means for following them.
A company can use any \ac{MDM} tool, curated app store or user agreement to enforce their policy.
\AppPAL~links the specification of the policy to its implementation, giving visibility into exactly how a policy is implemented and satisfied.

We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
Our AppPAL implementation can also be easily extended to support new types of policies.
It also gives us access to tooling we have developed to can check AppPAL policies for completeness and redundant statements.
For this work additional tooling was developed to help visualise policies and describe policies contents.
This was helpful for making comparisons between policies and checking our formalisation for mistakes.

%\paragraph{Contributions}

In summary, our work makes the following contributions:

\begin{itemize}
  \item We present a formalisation of five different BYOD policies in \AppPAL: a new instantiation of SecPAL for describing BYOD policies (\autoref{sec:idea}).
  \item Using our formalisation of the policies we make comparisons between the different policies. 
    Unlike previous work which looks at individual policies~\cite{armando_developing_2016}, our work looks at policies across a variety of domains (\autoref{ssec:byod-policies}).
  \item We identify that delegation and acknowledgements are an important aspect of BYOD policies that current MDM software does not look at (\autoref{sec:idioms}).
\end{itemize}

\subsection{Related Work}
\label{sec:related}
% 1-2 pages

Martinelli~\etal{}'s looks at creating a dynamic permissions manager, called UC-Droid.
Their tool can alter what an app's Android permissions are at run time based on policies~\cite{martinelli_enhancing_2016}.
The tool allows companies to reconfigure the apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
These kinds of policies are more configurable than the geofenced based policies existing MDM packages provide.
Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
The work allowed a company to verify the devices within their network and what servers and services they could access.
It also describes a mechanism for providing different users with different policies.

Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_bring_2013}.
Their tool allows companies to distribute apps through a secure app~store~\cite{armando_enabling_2014}.
The store ensures apps meet policies through a combination of static analysis and app rewriting with dynamic enforcement.
Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
Using their tool, they implemented parts of a NATO Communications and Information Agency policy relating to personal networks and data management~\cite{armando_developing_2016}.
Their work shows how the app-specific sections of a BYOD policy can be check and enforced using tools.
They did not look at where the checks or policies come from, however.

@@

Other work has looked at using app wrapping (where an app is recompiled to use guarded APIs) to enforce more general policies.
Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
App rewriting has the advantage that the device's underlying OS needn't be modified as the apps are changed at the source code level.
However app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.

Our approach taken with \AppPAL~is similar to work on safety cases.
A safety case is an argument made to say a system is acceptably safe to be used in a given scenario.
Industrial safety cases are often described in natural language, which can be ambiguous and unclear.
\ac{GSN}~\cite{tim_kelly_goal_2004} is one approach to make the safety cases explicit.
It is a graphical formal notation that lets engineers argue that a system is safe by linking top-level goals, to sub-goals to the arguments made for a system's safety.
Similarly work developing a formal language for specifying how medical staff should collaborate in a healthcare scenario~\cite{papapanagiotou_formal_2014} again helps clarify how roles are filled in a medical context, on the basis of staff and different healthcare providers.

It is interesting to examine how leading~\cite{rob_smith_magic_2016} MDM~tools such as IBM's MaaS360, MobileIron and \ac{BES}, enforce \emph{BYOD} policies.
Each of these tools supports enforcing and checking compliance policies. 
They do not, however, use policy languages to specify policies; rather they provide a limited number of checkboxes that admins can tweak (an excerpt of a policy from MaaS360 is shown in \autoref{fig:maaspolicy}).
These tools allow administrators to configure a device's settings and provision the devices with company apps.
Some support app wrapping, which enables them to encrypt app data locally, use a VPN within the app, or prevent apps not being used when the device isn't compliant.
Because the policies are inflexible and tightly coupled to the device's OS, intervention by an administrator is often required.
Whilst MDM software is good at configuring devices, selecting which policies to apply is typically a manual process performed by an administrator.
Removing blacklisted apps is a common feature, but the selection process of which apps to remove is manual.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/maas360-policy.png}
  \caption{Excerpt of a policy showing network settings from MaaS360.}
  \label{fig:maaspolicy}
\end{figure}

@@

As mobile devices have become more common in the workplace, BYOD policies have been written to help control the devices.
Part of their policies are prescriptive:  if you configure your device in this way, you will mitigate that threat.
The policies contain more than just configuration, however.
Consider this rule taken from the \emph{Security Policy for the use of handheld devices in corporate environments} by SANS~\cite{nicholas_r._c._guerin_security_2008}.

@@

A company could make use of an \ac{MDM} program.
Some can use geofencing to apply a policies in the area around a lab.
Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration.
It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
Can it judge using a policy if a location is restricted?
The rule also gives a security objective: \emph{prevent malicious users from sending proprietary pictures}.
The guidelines are given, however, for the case of a legitimate user using MMS or email.
It may not be sufficient to stop a sufficiently motivated \emph{malicious} user.

% To answer these questions we came up with \AppPAL{}: 
%   a formal language for linking policies to the tools used to implement them and distributing decisions about apps.
Our approach does not try to enforce the policy by checking the app's code for programming errors.
Rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration are made explicit.
This gives us greater clarity as to which tool is being trusted to implement what policy.
It allows us to see who is being trusted to make which decisions,
  and use automatic-tools to uncover problematic aspects of the policy~\cite{hallett_specifying_2016}.
Continuing with the example above, we can write this in \AppPAL~as:

@@

\section{Instantiating SecPAL}
\label{ssec:changes}

SecPAL was developed as a distributed access control language~\cite{becker_secpal:_2010}.
The SecPAL language is designed to be have a clear readable syntax, and intuitive semantics.
It is also designed to be extensible, which makes it ideal for extending to create new languages.
All SecPAL statements are \emph{said} by an explicit authority.
The authority can say a fact (that something is described by a predicate), a delegation (that someone else \emph{can-say} a fact), or a role assignment (that something \emph{can-act-as} something else).
This statement optionally contain conditional facts, and constraints that must be satisfied before the authority will say the statement.

@@

Facts of the \emph{must}-kind represent obligations, actions to complete if a particular scenario presents itself.
For these facts, we add a rule to check we perform the obligation.
This rule should be checked periodically to ensure compliance.
Our implementation contains tooling to generate these rules automatically, by parsing the policy.
\begin{lstlisting}
$\langle\text{speaker}\rangle$ says $\langle\text{subject}\rangle$ hasSatisfiedObligation$\langle\text{Action}\rangle$
  if $\langle\text{subject}\rangle$ must$\langle\text{Action}\rangle$,
     $\langle\text{subject}\rangle$ has$\langle\text{Action}\rangle$.
\end{lstlisting}

Facts using \emph{is} predicates give types to variables. 
\AppPAL~inherits from SecPAL's (and Datalog's) safety condition that the body of a statements must reference all the variables in the head.
This can lead to some \emph{boilerplate} code in policies that may obscure their meaning.
To simplify the policies, we add syntactic sugar for facts giving variables their type (\emph{variable \emph{is}Type}).
Variables in the head of the statement of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the condition.
The two statements shown below (taken from the SANS policy) are equivalent, however we feel the example on the right is easier to read.

\begin{minipage}{\linewidth}
  \begin{minipage}{0.5\linewidth}
\begin{lstlisting}
'company' says Device canConnectToAP(X)
  if X isOwnedByCompany,
     Device isDevice,
     X isAP.
\end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\linewidth}
\begin{lstlisting}
'company' says Device:D canConnectToAP(AP:X)
  if X isOwnedByCompany.
\end{lstlisting}
\end{minipage}
\end{minipage}

@@

We examined 5 policies and encoded them into \AppPAL{} looking for common idioms.
We selected these policies as they came from a variety of domains.
\begin{itemize}
  \item The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
This policy is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices.
Companies are expected to modify the document to suit their needs.
The policy is general; not specific to any particular industry, device, or country's legislation.
\item The second is taken from the \ac{HiMSS}~\cite{healthcare_information_and_management_systems_society_mobile_2012};
  a US non-profit company trying to improve healthcare through IT.
The \ac{HiMSS} policy is relatively short and contains concerns specific to healthcare scenarios. 
It is written as a contract the users agree to follow.
In contrast, every other policy we looked at is written as an organisation imposing rules on users they should follow to ensure compliance.
The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
\item The third is taken from a British hospital trust~\cite{kennington_mobiles_2014} and describes the BYOD scheme used in practice at the hospital.
\item Finally, we looked at two simpler policies from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
  These policies are simpler, and shorter than the other policies we looked at comprised of much more general rules.
\end{itemize}

\begin{table}\centering\footnotesize\sffamily
  \resizebox{\textwidth}{!}{%
    \def\myfntx{}
    \begin{tabular}{l c c c c c}
      \toprule
                                      & {SANS}       & {HiMSS}     & {NHS}       & {Edinburgh}                       & {Sirens}    \\
      \midrule
      Number of rules                 & 33           & 15          & 56          & 20                                & 25          \\
      \AppPAL~statements              & 71           & 21          & 58          & 10                                & 39          \\
      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large nmuber of rules that whilst marked as rules are infact just descriptions of the document.  All the policy rules that described restrictions or relationships were implemented in \AppPAL{}.} & 22 (88$\%$) \\
      \midrule
      Rules using Acknowledgement     & 2            & 10          & 11          & 1                                 & 6           \\
      Rules using Delegation          & 23           & 5           & 33          & 2                                 & 13          \\
      Rules describing a restriction  & 18           & 3           & 8           & 1                                 & 5           \\
      \midrule
      Principal Speaker               & company      & user        & nhs-trust   & records-management                & department  \\
    %Technical Speaker            & it-department & xyz-healh-system & it-department &                    & it-department \\
      \bottomrule                    \\
    \end{tabular}
  }
  \caption{Summary of the contents of each of the BYOD policies.}
  \label{tab:summary}
\end{table}

We summarise the policies in \autoref{tab:summary}.
Each policy contains a series of \emph{rules}, which we implemented by one or more \emph{\AppPAL~statements}.
The \emph{policy coverage} represents the number of rules that have an \AppPAL~description attached.

% Not all the rules in a policy require implementation.
% For example rule 9.1 in the NHS policy states:
% \begin{policyrule}{NHS}
%   Downloading of personal apps onto a corporately issued mobile device should be avoided where possible.
%   The Trust would not encourage staff members to download apps for personal use onto a corporately issued mobile device.
%   All staff are reminded that they must adhere to the guidance outlined in the Social Media Policy.
% \end{policyrule}
% The rule does not prohibit any action and doesn't require the staff to do anything different.
% It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
% In this case, however, since nothing is required even an acknowledgement seems excessive.

@@

\section{BYOD Idioms}
\label{sec:idioms}

When examining the policies, we noticed two particular idioms reappearing: acknowledgements and delegation.
We describe both idioms in greater detail, and show how they can be implemented in \AppPAL{}, below.
MDM tools and research have focussed so far on implementing restrictions on apps and devices~\cite{_ibm_????,armando_formal_2014,martinelli_enhancing_2016}.
Implementing these controls is a vital aspect of BYOD policies and all 5 of the policies we looked at had rules that described restrictions~(\autoref{tab:summary}).
Every policy also contained rules that required employees acknowledgements, however.
Only the SANS policy (which is configuration focussed) contained more rules that required restrictions than acknowledgements. 
All the policies contained more rules featuring delegation relationships than functionality restrictions.
%Restricting device functionality is tricky and important, but other aspects of BYOD policies are also worth attention.

@@

\AppPAL~requires an explicit speaker for each statement.
Speakers can delegate to others by making a statement about what they \emph{can-say}.
When translating the policies, the author of the policy is used as the primary speaker of the policy's rule (\autoref{tab:summary}).
For the \ac{HiMSS} policy, where the user states what they will do rather than the company stating what they must, the user is the primary speaker.
All the policies describe multiple entities that might make statements and delegate.
With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
The delegation might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.

@@

%\section{Comparison with Existing MDM Software}

% It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}, both leading MDM packages~\cite{rob_smith_magic_2016}.
% Each of these tools supports enforcing and checking compliance policies. 
% They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.

% Acknowledgements form a significant portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
% MaaS360 can associate users with devices and can assign them into groups.
% It can distribute electronic documents, but there are no means to track what policies a user should read and follow.
% Administrators could add users who have acknowledged a policy to a group.
% One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
